---
title: "DATA607 Project 1"
author: "ODONNELL"
date: "February 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, lets read in the text file as a large string
(for stringr, this worked better than reading it in as a dataframe)
```{r}
txtfile = readChar("tournamentinfo.txt", file.info("tournamentinfo.txt")$size)
```

Now, let's trim the first 360 characters from that text file
(The first 360 characters are the row headers and such)
```{r}
#using nchar to figure out total characters..
nchar(txtfile)
txtfile <- substr(txtfile, 360, 17836)
```


Next, lets extract each needed variable with stringr:
player_name
state
player_ID
total_points
pre_rating
```{r}
library(stringr)

#lets first extract all the players' names
player_name <- unlist(str_extract_all(txtfile, "[|][:space:][[:upper:]]{2,}[[:space:], [:upper:]]{2,}[[:upper:]$]"))
player_name <- unlist(str_sub(player_name, start = 3))
player_name
```

```{r}
#extract all state names
state <- unlist(str_extract_all(txtfile, "[[:upper:]]{2}[[:space:]]{1}[|]"))
state <- unlist(str_extract_all(state, "[[:upper:]]{2}"))
state
```

```{r}
#extract all players' ID numbers
player_ID <- unlist(str_extract_all(txtfile, "[[:digit:]]{1,2}[[:space:]]{1}[|]"))
player_ID <- unlist(str_extract_all(player_ID, "[[:digit:]]{1,2}"))
player_ID
```

```{r}
#extract total points
total_points <- unlist(str_extract_all(txtfile, "[|][0-6]{1}\\.{1}[[:digit:]]{1}"))
total_points <- unlist(str_extract_all(total_points, "[0-6]{1}\\.{1}[[:digit:]]{1}"))
total_points
```

```{r}
#extract pre-ratings
pre_rating <- unlist(str_extract_all(txtfile, "[[:digit:]P]{3,}[[:space:]]{0,4}[-]"))
pre_rating <- unlist(str_extract_all(pre_rating, "[[:digit:]]{3,4}"))
pre_rating
```

Now, lets enter all of the extracted fields into a dataframe!
```{r}
tourny_df = data.frame(player_ID, player_name, state, total_points, pre_rating)
head(tourny_df)
```

Unfortunately, all the columns in the df are type level
So, lets change the data type of some of the dataframe columns
```{r}
tourny_df <- transform(tourny_df, player_ID = as.numeric(as.character(player_ID)),
          pre_rating = as.numeric(as.character(pre_rating)),
          total_points = as.numeric(as.character(total_points)),
          player_name = as.character(player_name),
          state = as.character(state))
str(tourny_df)
```

Lastly, lets add in the Average Pre Rating of Opponents
The hardest step!
We will create a few new dataframes in the process..
```{r}
opponents <- unlist(str_extract_all(txtfile, "[|][0-6]{1}\\.{1}[[:digit:]]{1}[[:space:]]{1,}([|][[:upper:]]{1}[[:space:]]{1,}[[:digit:]]{0,2}){7}"))
opponents <- unlist(str_extract_all(opponents, "([|][[:upper:]]{1}[[:space:]]{1,}[[:digit:]]{0,2}){7}"))
opponents_df <- data.frame(unlist(str_split_fixed(opponents, "[|]", n=8)))
```

So, in the chunk above we created a dataframe with all opponents scores
But, we need to clean it up.
```{r}
#delete first column
opponents_df <- opponents_df[ , !names(opponents_df) %in% "X1"]
```

Check the data types in the opponents_df
```{r}
str(opponents_df)
```

Cool, now remove first two chars from each cell
```{r}
opponents_df$X2 <- substring(opponents_df$X2, 2)
opponents_df$X3 <- substring(opponents_df$X3, 2)
opponents_df$X4 <- substring(opponents_df$X4, 2)
opponents_df$X5 <- substring(opponents_df$X5, 2)
opponents_df$X6 <- substring(opponents_df$X6, 2)
opponents_df$X7 <- substring(opponents_df$X7, 2)
opponents_df$X8 <- substring(opponents_df$X8, 2)
```


Now, change the data type of all the dataframe columns
```{r}
opponents_df <- transform(opponents_df, X2 = as.numeric(as.character(X2)),
          X3 = as.numeric(as.character(X3)),
          X4 = as.numeric(as.character(X4)),
          X5 = as.numeric(as.character(X5)),
          X6 = as.numeric(as.character(X6)),
          X7 = as.numeric(as.character(X7)),
          X8 = as.numeric(as.character(X8)))
str(opponents_df)
```

Now, lets add a column that matches Player_ID to opponents_df
Reason being, then we can merge the two dataframes based on Player_ID
```{r}
opponents_df$player_ID = c(1:64)
```

OK, lets merge our two dataframes into a new, temporary dataframe
```{r}
merged_df <- merge(tourny_df, opponents_df, by = "player_ID", all = TRUE)
```

Now, lets use merge to do some vlookups
```{r}
merged_df$opp1 <- merge(merged_df[, c("X2", "pre_rating")], tourny_df, by.x = "X2", by.y = "player_ID", all.x = TRUE)
```

```{r}
merged_df$opp2 <- merge(merged_df[, c("X3", "pre_rating")], tourny_df, by.x = "X3", by.y = "player_ID", all.x = TRUE)
```

```{r}
merged_df$opp3 <- merge(merged_df[, c("X4", "pre_rating")], tourny_df, by.x = "X4", by.y = "player_ID", all.x = TRUE)
```

```{r}
merged_df$opp4 <- merge(merged_df[, c("X5", "pre_rating")], tourny_df, by.x = "X5", by.y = "player_ID", all.x = TRUE)
```

```{r}
merged_df$opp5 <- merge(merged_df[, c("X6", "pre_rating")], tourny_df, by.x = "X6", by.y = "player_ID", all.x = TRUE)
```

```{r}
merged_df$opp6 <- merge(merged_df[, c("X7", "pre_rating")], tourny_df, by.x = "X7", by.y = "player_ID", all.x = TRUE)
```

```{r}
merged_df$opp7 <- merge(merged_df[, c("X8", "pre_rating")], tourny_df, by.x = "X8", by.y = "player_ID", all.x = TRUE)
```

Create one last dataframe to calc average opponents pre rating
```{r}
opp_average <- data.frame(merged_df$opp1$pre_rating.x, merged_df$opp2$pre_rating.x,
                          merged_df$opp3$pre_rating.x, merged_df$opp4$pre_rating.x,
                          merged_df$opp5$pre_rating.x, merged_df$opp6$pre_rating.x,
                          merged_df$opp7$pre_rating.x)

opp_average$avg_opp_pre_rating <- rowMeans(opp_average)

opp_average$player_ID = c(1:64)

opp_average <- opp_average[, c("player_ID", "avg_opp_pre_rating")]
```

Lastly, add average opponents pre rating to our master dataframe:
tourny_df
```{r}
tourny_df <- merge(tourny_df, opp_average, by = "player_ID", all = TRUE)
head(tourny_df)
```

Finally, export our work as a CSV!
```{r}
write.csv(tourny_df, "tournament_info.csv")
```